---
title: "bt catch comparison"
output: html_document
date: "2025-03-17"
---

# ASIDE !!!!! 

taxonomic checks 
uncertain about: Lycodes brevipes, did not look very far down taxon list - best may be to first compare to trawl data and then evaluate any taxa that are not in that dataset. 
```{r}
catch <- read.csv("/home/kimberly.ledger/goa-bt23-mb/data/CATCH_AND_HAUL_DATA_GOA_BT_2023_Chordata.csv")
```

get full taxon lineage for catch
```{r}
library(taxize)

taxon_codes <- unique(catch$World.register.of.marine.species..WoRMS..taxonomic.serial.number)
taxon_codes_clean <- discard(taxon_codes, is.na)  # Remove NA values
taxonomy_list <- map(taxon_codes_clean, ~ classification(.x, db = "worms")[[1]])

#taxonomy_list2 <- taxonomy_list[!is.na(taxonomy_list)] #Remove NA values (replace `NA` check with `!is.na()`)

# Step 3: Convert list to a dataframe
taxonomy_df <- bind_rows(taxonomy_list, .id = "id") %>%
  mutate(taxon_id = taxon_codes_clean[as.numeric(id)]) %>%  # Map back to taxon_id
  filter(rank %in% c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")) %>%
  select(taxon_id, rank, name) %>%
  pivot_wider(names_from = rank, values_from = name) %>%
  select(taxon_id, Kingdom, Phylum, Class, Order, Family, Genus, Species)  # Ensure column order

# Step 4: Merge back with the original dataframe
catch_df <- catch %>%
  left_join(taxonomy_df, by = c("World.register.of.marine.species..WoRMS..taxonomic.serial.number" = "taxon_id") )
```

```{r}
catch_df %>%
  filter(is.na(Kingdom)) %>%
  select(Taxon.common.name, Taxon.scientific.name, World.register.of.marine.species..WoRMS..taxonomic.serial.number) %>%
  unique()
```

all these seem reasonable to exclude from catch 

catch summary (entire 2023 GOA BT data export of Chordata)
```{r}
catch_summary <- catch_df %>%
  filter(!is.na(World.register.of.marine.species..WoRMS..taxonomic.serial.number)) %>%
  group_by(Taxon.common.name, Taxon.scientific.name, Class, Order, Family, Genus, Species) %>%
  summarise(biomass_cpue = sum(Weight.CPUE..kg.km2.),
            n_cpue = sum(Number.CPUE..no.km2.)) %>%
  arrange(desc(biomass_cpue))
catch_summary 
```

```{r}
library(forcats)

catch_summary %>%
  filter(biomass_cpue > 10000) %>%
  ggplot(aes(x = fct_reorder(Taxon.common.name, biomass_cpue, .desc = T), y = biomass_cpue)) +
  geom_point(size = 3) +  # Adds points
  scale_y_log10() +
  labs(title = "Biomass - 2023 GOA Bottom Trawl Survey",
       x = "Species",
       y = "Biomass (kg)") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1))  # Rotates x-axis labels for readability
```




cross check eDNA assignments with catch data to see if there's any obvious misIDs of ASVs 

identify species in both eDNA and trawl
```{r}
both_lists <- intersect(taxon_table$taxon, catch_df$Taxon.scientific.name)
both_lists
```

identify species in eDNA but not trawl
```{r}
eDNA_only <- setdiff(taxon_table$taxon, catch_df$Taxon.scientific.name)
eDNA_only
```

NOTES:
Atheresthes: Atheresthes evermanni and Atheresthes stomias in trawl (Atheresthes stomias - Arrowtooth flounder VERY abundant) - uncollapsed taxonomy has both species 
Balaenoptera physalus: Fin whale - good thing that didn't end up in the net! 
Bathyagonus: Bathyagonus nigripinnis, Bathyagonus pentacanthus, and Bathyagonus alascanus in trawl - uncollapsed taxonomy has B. alascanus and B. infraspinatus - check if reference data is available for other species - if so, edna detections see to be B. alascanus 
Bathymaster: only Bathymaster signatus in trawl - uncollased tax has B. leurolepis and B. signatus
Careproctus: Careproctus sp. in trawl - uncollapsed tax are Careproctus phasma and Careproctus scottae
Cololabis saira: doesn't seem to be anywhere in the catch... but seems plausible to find  
Cryptacanthodes aleutensis:  Cryptacanthodes giganteus in trawl - check reference data availability 
Gadidae: pollock, pcod, saffron cod, and tomcod all in trawl - uncollapsed tax has pollock, pcod, and arctic cod. 
Gadus: pollock and pcod in trawl 
Icelinus borealis: this species not in trawl, but many other sculpins are 
Lepidopsetta: Lepidopsetta bilineata and Lepidopsetta polyxystra in trawl - uncollapsed tax has both spp
Megaptera novaeangliae: humpback whale
Oncorhynchus: chum, chinook, pink, and coho in trawl - uncollapsed tax has chum, sockeye and pink 
Oncorhynchus nerka: no sockeye in trawl... 
Phocoena phocoena: Harbour porpoise
Pleuronectidae: looking which species made up this assignment ... 
Psychrolutes paradoxus: not in trawl  (only shows up in one eDNA pcr rep)
Radulinus asprellus: not in trawl, but other sculpins are 
Ronquilus jordani: not in trawl 
Salvelinus: not in trawl - uncollapsed tax has S. alpinus and S. malma 
Sebastes: LOTS of species (mostly POP) - let's see what rkfish mb says 
Sebastes nigrocinctus: in range, surprising it's not in trawls this year?? also likely low confidence in mifish id - let's see what rkfish mb says 
Sebastolobus: mostly Sebastolobus alascanus in trawl but also Sebastolobus altivelis and Sebastolobus macrochir - let's see what rkfish mb says 
Stenobrachius leucopsarus: Stenobrachius sp. in trawl nets


## check out uncollapsed taxonomic assigns to see how these fit 

```{r}
taxonomy_uncollapsed <- read.csv("/home/kimberly.ledger/goa-bt23-mb/outputs/mifish_taxonomy_20250130_uncollapsed.csv") %>% 
  select(!X) %>%
  rename(ASV = qseqid)
```

explore pleuronectidae - it's the only non species or genus assignment with a good number of reads 
```{r}
pleuro <- taxonomy_uncollapsed %>%
  filter(taxon == "Pleuronectidae")
```

let me say: 
Pleuronectidae 1 = Isopsetta isolepis, Parophrys vetulus, Psettichthys melanostictus (ASV_018, ASV_151, ASV_201) - all three in trawl catch 
Pleuronectidae 2 = Glyptocephalus zachirus, Myzopsetta proboscidea (ASV_019, ASV_182) - lot of Glyptocephalus zachirus in catch - no Myzopsetta proboscidea
Pleuronectidae 3 = Hippoglossoides elassodon, Hippoglossoides robustus, Limanda aspera (ASV_037, ASV_047, ASV_202) - Hippoglossoides elassodon and Limanda aspera in catch 

```{r}
taxonomy_uncollapsed %>%
  filter(taxon == "Sebastolobus") %>%
  select(species) %>%
  unique()
``` 
